---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r interactive,echo=FALSE}
# setwd("/Data/CLASSHW/coursera/DataScienceSpecialization/05_repdata-008-hws/peer01/RepData_PeerAssessment1")
# library(knitr)
# knit2html("PA1_template.Rmd"); browseURL("PA1_template.html")
```

## What is mean total number of steps taken per day?
```{r totals and means}
data <- read.csv("activity/activity.csv")

complete <- data[complete.cases(data),]

days <- split(complete, complete$date, drop=TRUE)

totalStepsPerDay <- sapply(days, function(day) { return (sum(day$steps)) } )
```

```{r hist}
hist(totalStepsPerDay,
     main="Total steps per day",
     xlab="Steps")
```

## What is mean total number of steps taken per day?
The mean total number of steps taken per day is
`r as.integer(mean(totalStepsPerDay))`.
The median total number of steps taken per day is
`r as.integer(median(totalStepsPerDay))`.


## What is the average daily activity pattern?

```{r daily activity pattern}
intervals <- split(complete, complete$interval)

meanStepsPerInterval <- sapply(intervals, function(i) { return (mean(i$steps)) } )
```


```{r plot}
plot(meanStepsPerInterval,
     type="l",
     main="Daily Activity Pattern",
     ylab="Steps",
     xlab="Interval")

highestInterval = which.max(meanStepsPerInterval)
highestValue = meanStepsPerInterval[highestInterval]
```


The highest mean number of steps (of value `r highestValue`) occurs at five-minute interval `r highestInterval`.

## Inputing missing values
The total number of rows with missing values is
`r nrow(data) - nrow(complete)`.

```{r replaceMissing}
# Estimate the missing values
estimData = data
for (r in 1:nrow(estimData)) {
  if (is.na(estimData[r, 'steps'])) {
    interval = estimData[r, 'interval']
    index = which(names(meanStepsPerInterval) == as.character(interval))
    replacement = meanStepsPerInterval[index]
    estimData[r, 'steps'] = replacement
  }
}
```

```{r estimDays}
estimDays <- split(estimData, estimData$date, drop=TRUE)

estimTotalStepsPerDay <- sapply(estimDays, function(day) { return (sum(day$steps)) } )
```

```{r estimHist}
hist(estimTotalStepsPerDay,
     main="Total steps per day (with missing values estimated)",
     xlab="Steps")
```

With missing values filled in with estimates,
the mean total number of steps taken per day is
`r as.integer(mean(estimTotalStepsPerDay))`.
and the median total number of steps taken per day is
`r as.integer(median(estimTotalStepsPerDay))`.

```{r impact}
# What is the impact of adding estimates for the missing data?
# Returns: the percentage difference (roughly speaking)
Impact <- function (before, after) {
  return (200 * (before - after) / (before + after))
}

impactMean = Impact(mean(totalStepsPerDay), mean(estimTotalStepsPerDay))

impactMedian = Impact(median(totalStepsPerDay), median(estimTotalStepsPerDay))
```

The mean   changed by `r impactMean` percent.
The meidan changed by `r impactMedian` percent.
So we conclude that estimating the missing values did not significantly affect the results.

## Are there differences in activity patterns between weekdays and weekends?
